---
title: "STAT 600: Homework 1"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
author: "Hyungjoon Kim, Colorado State University"
output: pdf_document
---

```{r setup, echo=FALSE}
knitr::opts_knit$set(root.dir = './',
                     echo = TRUE,
                     message = FALSE,
                     warnings = FALSE)
```

# Problem 1, 2
I made an `R` package named `SimpLin`, and uploaded to my own `github` repository. You may check the structure and code of the package [**this link**](https://github.com/hjkim8987/STAT600-2024S/tree/main/HW1/SimpLin). You can also install that package including the vignette using the following code:

```{r, echo=TRUE, eval=FALSE}
devtools::install_github("hjkim8987/STAT600-2024S/HW1/SimpLin",
                         build_vignettes = TRUE)
```

# Problem 3  
I compared the performance of `lm` with `SimpLinR` using `rbenchmark` library. Each simulation data has $n=100$ observations, and the number of simulations is also $100$. I used $7$ cores for the parallelization, and `SimpLinR` is almost twice faster than the built-in `lm` function. Even though the result of `lm` function contains much more information than `SimpLinR`, using `Rcpp` makes a huge improvement of speed.  

```{r, echo=TRUE, eval=TRUE}
library(SimpLin)

library(foreach)
library(doParallel)

library(rbenchmark)

n_iter <- 100
n_obs <- 100

n_cores <- detectCores()

registerDoParallel(n_cores - 1)

benchmark(
  "SimpLinR" = {
    set.seed(600)
    
    ans <- foreach(i = 1:n_iter, .combine = "c") %dopar% {
      x <- rnorm(n_obs)
      eps <- rnorm(n_obs)
      y <- 1 - x + eps
      
      return(as.list(SimpLinR(x, y)))
    }
  },
  "lm" = {
    set.seed(600)
    
    ans <- foreach(i = 1:n_iter, .combine = "c") %dopar% {
      x <- rnorm(n_obs)
      eps <- rnorm(n_obs)
      y <- 1 - x + eps
      
      return(list(lm(y ~ x)))
    }
  },
  
  replications = 100,
  columns = c("test", "replications", "elapsed",
              "relative", "user.self", "sys.self")
)

stopImplicitCluster()
```

# Problem 4  
I attached three tables which summarize the simulation. According to the simulation, the built-in `lm` function and `SimpLin` give the exactly same results. They both have the same estimates, confidence intervals, bias, and so on. However, despite the simulation was implemented using the same-size data, `lm` took much more time for the fit.

```{r, echo=TRUE}
n_iter <- 100
n_obs <- 100

set.seed(600)

elapsed_time <- matrix(0, nrow = n_iter, ncol = 2)
bias <- matrix(0, nrow = n_iter, ncol = 4)
coverage <- matrix(0, nrow = n_iter, ncol = 4)
mse <- matrix(0, nrow = n_iter, ncol = 2)
pmse <- matrix(0, nrow = n_iter, ncol = 2)
beta_est <- matrix(0, nrow = n_iter, ncol = 4)

true_beta <- c(1, -1)
degrees_of_freedom <- n_obs - length(true_beta)
for(i in 1:n_iter){
  x <- rnorm(n_obs)
  eps <- rnorm(n_obs)
  y_true <- true_beta[1] + true_beta[2] * x
  y <- y_true + eps
  
  t1 <- Sys.time()
  fit_lm <- lm(y ~ x)
  t2 <- Sys.time()
  fit_rcpp <- SimpLinR(x, y)
  t3 <- Sys.time()
  
  elapsed_time[i, ] <- c(t2 - t1, t3 - t2)
  bias[i, ] <- c(true_beta - fit_lm$coefficients,
                 true_beta - c(fit_rcpp$coefficients))
  coverage[i, ] <- c((confint(fit_lm)[1, 1] < true_beta[1]) * 
                       (confint(fit_lm)[1, 2] > true_beta[1]),
                     (confint(fit_lm)[2, 1] < true_beta[2]) * 
                       (confint(fit_lm)[2, 2] > true_beta[2]),
                     (fit_rcpp$confidence.intervals[1, 1] < true_beta[1]) * 
                       (fit_rcpp$confidence.intervals[1, 2] > true_beta[1]),
                     (fit_rcpp$confidence.intervals[2, 1] < true_beta[2]) * 
                       (fit_rcpp$confidence.intervals[2, 2] > true_beta[2]))
  mse[i, ] <- c((t(fit_lm$residuals) %*% fit_lm$residuals)[1, 1] / 
                  degrees_of_freedom,
                (t(fit_rcpp$residuals) %*% fit_rcpp$residuals)[1, 1] /
                  degrees_of_freedom)
  pmse[i, ] <- c((t(y_true - fit_lm$fitted.values) %*% 
                    (y_true - fit_lm$fitted.values))[1, 1] / n_obs,
                 (t(y_true - fit_rcpp$fitted.values) %*% 
                    (y_true - fit_rcpp$fitted.values))[1, 1] / n_obs)
}
```  
```{r, echo=FALSE}
res1 <- round(cbind(colMeans(elapsed_time),
                    colMeans(mse),
                    colMeans(pmse)), 6)
colnames(res1) <- c("Elapsed Time", "MSE", "PMSE")
rownames(res1) <- c("`lm`", "`SimpLin`")
knitr::kable(res1,
             caption = "Average Elapsed Time, MSE, and PMSE")

res2 <- matrix(colMeans(coverage), nrow = 2)
colnames(res2) <- c("$\\beta_0$", "$\\beta_1$")
rownames(res2) <- c("`lm`", "`SimpLin`")
knitr::kable(res2,
             caption = "Average Coverage Rate of 95% Confidence Interval")

res3 <- matrix(colMeans(bias), nrow = 2, byrow = TRUE)
colnames(res3) <- c("$\\beta_0$", "$\\beta_1$")
rownames(res3) <- c("`lm`", "`SimpLin`")
knitr::kable(res3,
             caption = "Average Bias of the Estimation")
```
